---
title: "DSlecture10"
author: "Hening cui"
date: "2021/10/12"
output: github_document
---

```{r setup, message = FALSE}
library(tidyverse)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)
options(
  ggplot2.continuous.color = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
```

## group_by

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  ungroup(month)
```
## counting things

count month observation

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(n_obs = n())
```

we can use `count()`

```{r}
weather_df %>% 
  count(month, name = "n_obs")
```

never use bases r table function

```{r,eval = FALSE}
weather_df %>% 
  pull(month) %>% 
  table()
```
not df, cannot use

other counting function

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarise(
    n_obs = n(),
    n_days = n_distinct(date)
  )
```

## a digression on 2x2 table

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5  ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
    )
  ) %>% 
  janitor::tabyl(name, cold)
 # group_by(name, cold) %>% 
  #summarize(count = n()) %>% 

```


## general summarize

you can do lots summary

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarise(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE)
  )
```

this is a data frame

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarise(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_line()
```

suppose we want to summarize column
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarise(
    across(prcp:tmin, mean)
  )
```


reminder , some result ease to read in other format

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarise(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name, 
    values_from = mean_tmax
  ) %>% 
  knitr::kable(digits = 3)
```


## group_by and mutate


```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax
  ) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) +
  geom_point()
```

what about window

ranking

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  mutate(temp_rank = min_rank(desc(tmax)) 
  ) %>% 
  filter(temp_rank == 1)
```
  
lag whole column shift one 
could use to calculate change

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(lag_temp = lag(tmax))
```

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(lag_temp = tmax - lag(tmax)) %>% 
  summarise(
    temp_change_max = max(lag_temp, na.rm = TRUE),
    temp_CHANGE_sd = sd(lag_temp, na.rm = TRUE) 
    )
```

## quick note

summarize only get you so far


