---
title: "Dslecture10assessment"
author: "Hening cui"
date: "2021/10/12"
output: github_document
---

```{r setup, message = FALSE}
library(tidyverse)
library(haven)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)
options(
  ggplot2.continuous.color = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

In the PULSE data, the primary outcome is BDI score; it’s observed over follow-up visits, and we might ask if the typical BDI score values are roughly similar at each. Try to write a code chunk that imports, cleans, and summarizes the PULSE data to examine the mean and median at each visit. Export the results of this in a reader-friendly format.
assessment 1
```{r}
pulse_df =
  read_sas("./data/public_pulse_data.sas7bdat") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  group_by(id) %>% 
  summarise(
    mean_BDI = mean(bdi_score_bl:bdi_score_12m)
  )
```

```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)

pulse_data %>% 
  group_by(visit) %>% 
  summarize(
    mean_bdi = mean(bdi, na.rm = TRUE),
    median_bdi = median(bdi, na.rm = TRUE)) %>% 
  knitr::kable(digits = 3)
```


assessment 2

In the FAS data, there are several outcomes of interest; for now, focus on post-natal day on which a pup is able to pivot. Two predictors of interest are the dose level and the day of treatment. Produce a reader-friendly table that quantifies the possible associations between dose, day of treatment, and the ability to pivot.

```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./data/FAS_litters.csv") %>%
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_data = left_join(pup_data, litter_data, by = "litter_number") 

fas_data %>% 
  group_by(dose, day_of_tx) %>%
  drop_na(dose) %>% 
  summarize(mean_pivot = mean(pd_pivot, na.rm = TRUE)) %>%
  pivot_wider(
    names_from = dose,
    values_from = mean_pivot
  ) %>% 
  knitr::kable(digits = 3)
  
```


